---
title: "SuSiE RSS with QC"
output: html_document
date: "2023-12-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

We demonstrates how we use SuSiE RSS wrapper implemented in `pecotmr` package to fine-map using summary statistics and reference LD panel data precomputed and available as resources.

To load the `pecotmr` package,

```{r load-pkgs}
library(pecotmr)
```

## Load summary statistics and corresponding LD matrix

For given region of interest, we load the LD matrix and corresponding summary statistics data, where the summary statistics data should be a data frame with the columns "chr" and "pos":

```{r}
library(data.table)
sumstats = fread("ADGWAS2022.chr1.sumstat.tsv")
LD_meta_data = data.frame(chrom = "chr1",start = 50179506, end = 54277735, path = "ROSMAP_NIA_WGS.leftnorm.filtered.filtered.chr1_50179506_54277735.float16.txt.xz")
region = c("chr1", 51800000, 53927212)
```

```{r}
LD_data = load_LD_matrix(LD_meta_data, region, sumstats)
```

The output LD_data is a list contains the `LD` and `variants_df`, where the `variants_df` is a data frame with the columns "chr","variants","GD(genetic distance)"," pos","A1" and "A2" in a format of bim file. The row and column names of LD are identical to the elements of the `variants` in the data frame `variants_df`. The `variants_df` will be the input of `allele_qc` as the reference panel data.

## Summary statistics QC

Match summary statistics data and `variants_df` by ("chr", "A1", "A2" and "pos"), accounting for possible strand flips and major/minor allele flips (opposite effects and zscores). We need to ensure that the format of summary statistics data with the columns "chr", "pos", "A1" and "A2".

```{r}
allele_fip = allele_qc(sumstats, LD.data$variants_df, match.min.prop=0.2, remove_dups=TRUE, flip=TRUE, remove=TRUE)
```

The output `allele_flip` will be the summary statistics after allele flip.

Because we match sumstats and `LD.data$variants_df` by chrom and position after ``allele_qc` step, we need to use the variants of output to extract the `LD.data$LD`.

```{r}
library(dplyr)
allele_flip = allele_flip%>%mutate(variant_allele_flip =         paste(chr,pos,A1.sumstats,A2.sumstats,sep=":"))
LD_extract = LD.data$LD[allele_flip$variant_allele_flip,allele_flip$variant_allele_flip]
```

## run SuSiE RSS:

```{r}
res = susie_rss(allele_flip$z, LD_extract)
```